Doubly-Robust Inference in R using \\texttt{drtmle}

### 2023-01-04 paper/reviews/response.md

The article does a good job of motivating and describing the underlying methodology well for readers familiar with doubly robust estimators of the ATE, but not with the doubly robust inference literature. Overall, the manuscript is also very clearly written, and the package seems to have a nice design. Specific comments are detailed below.

The workhorse functions (drtmle and adaptive_iptw) allow for missingness in treatment (A) and outcome (Y) values (e.g., with DeltaA and DeltaY control parameters within ‘glm_g’). This bivariate missingness setting doesn’t appear (unless I’m missing something) to have been formally considered in either van der Laan (2014) or Benkeser et al. (2017). I am concerned whether such missingness may lead to non-identifiability of the counterfactual mean—or “treatment-specific marginal mean” as referred to by the authors—functionals, without further assumptions. Are the authors implicitly invoking concrete missingness assumptions under which the implemented approaches target the original functional of interest (e.g., something like Delta_A x Delta_Y independent of (A, Y) given W)? Or are the methods simply standard, or used for convenience? It would enhance the paper if this issue were clarified.

_Response_: Thanks for the suggestion. We have added the formal assumptions needed for identification to that section (copied here):

if we define $\Delta_A$ and $\Delta_Y$ as indicators that $A$ and $Y$ are observed, respectively, then we can still perform valid inference on $E[Y(a)]$ under the assumptions that (i) $Y(a) \perp \Delta_A \mid W$, (ii) $Y(a) \perp A \mid \Delta_A = 1, W$, and
(iii) $Y(a) \perp \Delta_Y \mid \Delta_A = 1, A = a, W$.

@@

Related to point 1., why is missingness in the covariates (W) not considered? I am aware that this would be a very challenging scenario methodologically (and this is a good enough reason to not deal with it), but I am just curious whether there is something fundamentally different about W in this setting which precludes consideration of it being partially missing. Any guidance on what a practitioner should do if W is partially missing?

_Response_: Thanks for the suggestion. We have added a paragraph on missingness in $W$, copied below.

> In practice it is common to also have missingness in $W$. In this
situation, there are varying assumptions under which one can still identify and 
estimate $E[Y(a)]$. The only native solution in `drtmle` currently is to augment $W$
with columns of missingness indicators for each component of $W$. For example, if
variable $W_1$ is observed, we could define a new indicator $\Delta_{W,1} = 1$ and 
$\Delta_{W,1} = 0$ otherwise. For observations with $\Delta_{W,1} = 0$, we could use
some (possibly very simple) imputation strategy to fill in a value for $W_1$. Candidate
learners for the OR and PS may then include appropriate cross-products between $W_1$ 
and $\Delta_{W,1}$. Other learners may opt to treat the imputed value of $W_1$ as 
the true value and ignore $\Delta_{W,1}$. Super learner can be used to help determine
the modeling strategy that leads to the optimal model fit conditional on $W_1$ and
$\Delta_{W,1}$. It is important to note that this approach is only valid for drawing 
causal inference under the assumption that $Y(a)$ is independent of $A$ conditional on
the partially observed variables in $W$ and their missingness indicators. This assumption
may not be plausible depending on the form of missingness of $W$. Other approaches for 
handling missing data, e.g., multiple imputation, could in theory be coupled with `drtmle`, 
though such applications have not been studied empirically and are not available natively
in the `drtmle` package.

### 2023-01-04 paper/using_drtmle_obstudies.Rmd
finalize revision

The `drtmle()` function supports missing data in `A` and `Y`. Such missing data
are common in practice and, if ignored, can bias estimates of the marginal
means of interest. The estimators previously discussed can be modified to allow
such missingness processes to depend on observed covariates. That is, if we 
define $\Delta_A$ and $\Delta_Y$ as indicators that $A$ and $Y$ are observed, respectively,
then we can still perform valid inference on $\psi_0$ under the assumptions that 
(i) $Y(a) \perp \Delta_A \mid W$, (ii) $Y(a) \perp A \mid \Delta_A = 1, W$, and
(iii) $Y(a) \perp \Delta_Y \mid \Delta_A = 1, A = a, W$. 
In this case, we can
redefine the OR to be $\overline{Q}_0(a,w) \coloneqq E_0(\Delta_Y Y \mid \Delta_A
A = a, \Delta_A = 1, \Delta_Y = 1, W = w)$ and similarly redefine the PS to be
\begin{align}
  g_0(a \mid w) &= Pr_0(\Delta_A = 1, \Delta_A A = a, \Delta_Y = 1 \mid W = w)
  \nonumber \\
  &= Pr_0(\Delta_A = 1 \mid W) Pr_0(\Delta_A A = a \mid \Delta_A = 1, W = w)
  (\#eq:modPS) \\
  &\hspace{1.5in} Pr_0(\Delta_Y = 1 \mid \Delta_A = 1, \Delta_A A = a, W = w)
   \nonumber.
\end{align}

@@

\textbf{Remark: } In practice it is common to also have missingness in $W$. In this
situation, there are varying assumptions under which one can still identify and 
estimate $\psi_0$. The only native solution in `drtmle` currently is to augment $W$
with columns of missingness indicators for each component of $W$. For example, if
variable $W_1$ is observed, we could define a new indicator $\Delta_{W,1} = 1$ and 
$\Delta_{W,1} = 0$ otherwise. For observations with $\Delta_{W,1} = 0$, we could use
some (possibly very simple) imputation strategy to fill in a value for $W_1$. Candidate
learners for the OR and PS may then include appropriate cross-products between $W_1$ 
and $\Delta_{W,1}$. Other learners may opt to treat the imputed value of $W_1$ as 
the true value and ignore $\Delta_{W,1}$. Super learner can be used to help determine
the modeling strategy that leads to the optimal model fit conditional on $W_1$ and
$\Delta_{W,1}$. It is important to note that this approach is only valid for drawing 
causal inference under the assumption that $Y(a) \perp A \mid \Delta_W W$, which may 
not be plausible depending on the form of missingness of $W$. Other approaches for handling
missing data, e.g., multiple imputation, could in theory be coupled with `drtmle`, 
though such applications have not been studied empirically and are not available natively
in the `drtmle` package.

>>>>

The `drtmle()` function supports missing data in `A` and `Y`. Such missing data
are common in practice and, if ignored, can bias estimates of the marginal
means of interest. The estimators previously discussed can be modified to allow
such missingness processes to depend on observed covariates. That is, if we 
define $\Delta_A$ and $\Delta_Y$ as indicators that $A$ and $Y$ are observed, respectively,
then we can still perform valid inference on $E[Y(a)]$ under the assumptions that 
(i) $Y(a) \perp \Delta_A \mid W$, (ii) $Y(a) \perp A \mid \Delta_A = 1, W$, and
(iii) $Y(a) \perp \Delta_Y \mid \Delta_A = 1, A = a, W$. 
In this case, we can
redefine the OR to be $\overline{Q}_0(a,w) \coloneqq E_0(\Delta_Y Y \mid \Delta_A
A = a, \Delta_A = 1, \Delta_Y = 1, W = w)$ and similarly redefine the PS to be
\begin{align}
  g_0(a \mid w) &= Pr_0(\Delta_A = 1, \Delta_A A = a, \Delta_Y = 1 \mid W = w)
  \nonumber \\
  &= Pr_0(\Delta_A = 1 \mid W) Pr_0(\Delta_A A = a \mid \Delta_A = 1, W = w)
  (\#eq:modPS) \\
  &\hspace{1.5in} Pr_0(\Delta_Y = 1 \mid \Delta_A = 1, \Delta_A A = a, W = w)
   \nonumber.
\end{align}

@@

\textbf{Remark: } In practice it is common to also have missingness in $W$. In this
situation, there are varying assumptions under which one can still identify and 
estimate $E[Y(a)]$. The only native solution in `drtmle` currently is to augment $W$
with columns of missingness indicators for each component of $W$. For example, if
variable $W_1$ is observed, we could define a new indicator $\Delta_{W,1} = 1$ and 
$\Delta_{W,1} = 0$ otherwise. For observations with $\Delta_{W,1} = 0$, we could use
some (possibly very simple) imputation strategy to fill in a value for $W_1$. Candidate
learners for the OR and PS may then include appropriate cross-products between $W_1$ 
and $\Delta_{W,1}$. Other learners may opt to treat the imputed value of $W_1$ as 
the true value and ignore $\Delta_{W,1}$. Super learner can be used to help determine
the modeling strategy that leads to the optimal model fit conditional on $W_1$ and
$\Delta_{W,1}$. It is important to note that this approach is only valid for drawing 
causal inference under the assumption that $Y(a)$ is independent of $A$ conditional on
the partially observed variables in $W$ and their missingness indicators. This assumption
may not be plausible depending on the form of missingness of $W$. Other approaches for 
handling missing data, e.g., multiple imputation, could in theory be coupled with `drtmle`, 
though such applications have not been studied empirically and are not available natively
in the `drtmle` package.
